---
title: "incarceration.Rmd"
author: "Anica Na"
date: "11/23/2021"
output: html_document
---

```{r setup, include=FALSE}
source("analysis.R")
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)
options(scipen=999)
```

## Jailing Population Analysis - Anica Na

Welcome to my website report analyzing the incarceration trends in the United States over different races. Through my report I specifically focused on one race in particular and analyzed white people's jailing rates in comparison to other races. As well as in comparison to the metropolitan population and also compiled a map displaying the concentration jailed caucasians all throughout America. To accomplish this, I first analyzed which county in America had the highest jail admission rate in the most recent year of our data, which was 2018. The county was _`r most_admission_county`_ and the count of individuals admitted was __*`r count_most_admission`*__. And out of that large value the amount of white people jailed in  _`r most_admission_county`_ was __*`r count_la_white`*__. Telling us that white cellmates were __*`r percent_white`*__ of the total jail admissions in Los Angeles during 2018. And then to check whether this data was only applicable to recent years I then analyzed for the county with the most incarcerations to count since 1970, and the answer was _`r max_incarcerated_county`_ again. The year this took place was in _`r max_incarcerated_year`_ and the amount of individuals incarcerated was _`r count_max_incarcerated`_. Which is practically 3x the count from 2018. Finally, I analyzed for the counties in America with the highest percentage of white incarcerated inmates in 2018. Resulting in the _`r highest_county_white`_ data vector which tells us that 75 counties in the United States had a 100% white inmate population incarcerated. Informing me that perhaps due to the much large population index of caucasian individuals in comparison to the other races that they are very likely the most incarcerated race in America.

## Visual Representations

# Trend over Time
```{r echo=FALSE}
race_in_jail
```
My first chart above displays all of the races and the number of each incarcerated throughout the years. Through this we can easily view that white people and black people were the two races most incarcerated with black individuals being higher than white for a period. However, from the late 1990s and since then white people have been the most incarcerated. This is then followed by black, latinx, native american, and the asian american race has the lowest number of incarcerated individuals. I did expect the top three races to be black, white, and latinx but I was slightly surprised seeing how white was the highest. This is most likely due to recent media centering around the crisis of racism and bringing to light the overwhelming injustice of unfairly incarcerated african american individuals. However, I do believe that if the length of prison sentences or the ratio of overturned convictions was looked into the african american line would be much higher than the other races.

# Variable Comparison
```{r echo=FALSE}
metro_jail_pop
```
My next chart is a scatterplot that looks at the 
# Map
```{r echo=FALSE}
white_pop_map
```
