---
title: "US Incarceration Rates based on Race"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("analysis.R")
```

## Jailing Population Analysis - Anica Na

Welcome to my website report analyzing the incarceration trends in the United States over different races.Through my report I specifically focused on one race in particular and analyzed white people's jailing rates in comparison to other races. As well as in comparison to the metropolitan population and also compiled a map displaying the concentration jailed caucasians all throughout America. To accomplish this, I first analyzed which county in America had the highest jail admission rate in the most recent year of our data, which was 2018. The county was _`r most_admission_county`_ and the count of individuals admitted was __*`r count_most_admission`*__. And out of that large value the amount of white people jailed in  _`r most_admission_county`_ was __*`r count_la_white`*__. Telling us that white cellmates were __*`r percent_white`*__ of the total jail admissions in Los Angeles during 2018. And then to check whether this data was only applicable to recent years I then analyzed for the county with the most incarcerations to count since 1970, and the answer was _`r max_incarcerated_county`_ again. The year this took place was in _`r max_incarcerated_year`_ and the amount of individuals incarcerated was _`r count_max_incarcerated`_. Which is practically 3x the count from 2018. Finally, I analyzed for the counties in America with the highest percentage of white incarcerated inmates in 2018. Resulting in the data frame called, __*highest_county_white*__ which cannot be displayed because it tells us that 75 counties in the United States had a 100% white inmate population. Leading me to predict that perhaps due to the much larger population index of caucasian individuals in comparison to other races, that they are very likely the most incarcerated race in America.

## Visual Representations

# Trend over Time
```{r echo=F, warning=F, error=F, message=F}
print(race_in_jail)
```

My first chart above displays all of the races and the number of each incarcerated throughout the years. \
Through this we can easily view that white people and black people were the two races most incarcerated with black individuals being higher than white for a period. \
However, from the late 1990s and since then white people have been the most incarcerated. \
This is then followed by black, latinx, native american, and the asian american race has the lowest number of incarcerated individuals. \
I did expect the top three races to be black, white, and latinx but I was slightly surprised seeing how white was the highest. \
This is most likely due to recent media centering around the crisis of racism and bringing to light the overwhelming injustice of unfairly incarcerated african american individuals. \
However, I do believe that if the length of prison sentences or the ratio of overturned convictions was looked into the african american line would be much higher than the other races.

# Variable Comparison
```{r echo=F, warning=F, error=F, message=F}
print(metro_jail_pop)
```

When looking at The Vera Institute [Codebook](https://github.com/vera-institute/incarceration_trends/blob/master/incarceration_trends-Codebook.pdf?raw=true), metropolitan area refers to the [Core-Based Statistical Area (CBSA)](https://www.census.gov/topics/housing/housing-patterns/about/core-based-statistical-areas.html), which is the area of a metropolitan or micropolitan in a county. My variable comparison chart above shows the relationship between the metropolitan size and the number of white individuals within the jail population. While there are no clear, blatant patterns in the chart, this also means that the size of the city does not influence the number of incarcerated white people. This, however, does not imply that mass incarceration does not exist in these cities, this chart simply demonstrates that city size is not an obvious factor. This graph is included in the report to show that regardless of the metropolitan size, some cities have more than 1000 incarcerated whites and vice versa, there are cities with 0 incarcerated whites, with those cities all being different sizes as well. 

# Map
```{r echo=F, warning=F, error=F, message=F}
white_pop_map
```

My map above depicts the percentage of incarcerated whites in each county all across America. While a large majority of the map and the counties are filled in with varying shades of blue, some noticeable patters are a cluster of white counties near the center north of our chart. This implies that in those counties there is an extremely low or non-existent percentage of white people in jail there. This could be due to an extremely low caucasian population or even a very low general population as well in those areas. This map was created to show not only the percentage of jailed whites but also show the power of proportion as visualizing multiple, large numbers would be very hard. However, by using percentages, it is much clearer to pick out which counties have low, high, or even 100% of inmates being of the white race.